\documentclass[10pt]{article}

\usepackage[letterpaper]{geometry}
\usepackage{enumerate,verbatim,mdwlist}
\usepackage{fancyhdr}
\usepackage{linguex}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}

%Packages needed for trees
\usepackage{amsfonts,amsmath,amssymb}
\usepackage[varg]{txfonts}
\usepackage{qtree}

%% Margin Setting
\geometry{hmargin={.5in,.5in},vmargin={1in,1in}}
\setlength{\parindent}{0.0in}
\setlength{\parskip}{3mm}
\setlength{\tabcolsep}{10pt}
\setlength{\arraycolsep}{10pt}

%% Header
\setlength{\headheight}{23pt}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[L]{Phil 101, f14}
\fancyhead[C]{Homework \# 2}
\fancyhead[R]{Due: Thursday October 2, 2014 \\ Point total: 20 points}

\begin{document}

\textbf{Name:}\underline{\hspace{2in}}

\paragraph{Directions:} For each of the passages below, determine what, if any, informal fallacy the argument commits.  It can be any of the fallacies that we discussed in class, but some passages may not commit a fallacy at all.  You only need to write down one of the names for the appropriate fallacy, but if you provide a short explanation of your answer, I will take it into account in assigning partial credit.  You may find it helpful to reconstruct the argument in premise/conclusion form to better bring out the fallacy involved. (\textbf{2 points each})

\begin{enumerate}
  \setlength\itemsep{45pt}

  \item ``There's no reason for you to help out at the soup kitchen.  After all, the hungry people there are a part of the total problem of world hunger.  And world hunger is a huge problem that no person can solve on their own. So, your efforts at the soup kitchen won't do any good.''
  
  \item ``The teacher asked me to pass the exam out to all the students, which I did.  Since I passed the exam, I think I deserve an A in the course.''
  
  \item ``The governor says we need to alleviate overcrowding in prisons by releasing some of the prisoners.  But prisons are there for a reason.  There are serious criminals out there that make all of us afraid to be out on the streets.  Basically, the governor wants to turn us into a crime state. His plan is crazy!''
  
  \item ``Same-sex marriage should never be allowed in this state. If we allow gays to marry each other, then in no time uncles will marry their nephews and nieces. Then fathers will marry their daughters, mothers will marry their sons, and brothers will marry their sisters. Before
long, pet owners will marry their dogs and cats, and this will lead to the complete destruction of civilized life.''

  \item ``These environmentalists have gone overboard. Who are these crazies who dictate what we can do with our own property? This great nation was founded in the spirit of capitalism, and free enterprise! The sacred right of private property is grounded in the commandments of almighty God! Down with the nature-nuts! Away with the eco-maniacs!''
  
  \item ``Libertarians are insensitive to the plight of the poor. This is made evident by their plan to abolish the minimum wage. They want to do this because they want to eliminate all restrictions on hiring the poor. And of course this is true because Libertarians just don't care about the poor.''
  
  \item ``Senators Harrison and Beaney are both republicans from states with small populations.  They're both known outdoorsmen, and they have records of sticking mostly by the party line.  Harrison supports the concealed weapon bill.  So, Beaney is probably in favor of it as well.''
  
  %\item ``Max argues that we should upgrade our firefighting equipment by purchasing a fleet of new trucks with all the bells and whistles. But it’s only natural that he should argue this way. Max is a firefighter himself, and he just wants a fancy new truck to ride around in.''
  
  \item ``A spokesman for the government has argued that our insistence on a twelve percent increase in pay for steel workers is inflationary. But the government's own actions are far more inflationary than ours. The government just gave a fifteen percent increase to every single employee of the giant federal bureaucracy.''
  
  \item ``Clearly the moral standards of our youth have decayed. Two doctors at Central Hospital appeared recently on a talk show, and both were convinced that the youth of today have no morals at all.''
  
  \item `` Most elderly people who are hospitalized with cancer eventually die from the disease. Thus, if an elderly person wants to recover from cancer he or she must, at all costs, refuse hospitalization.''
  
\end{enumerate}  

\vspace{1in}
  
\paragraph{Bonus:}  The fallacies of \textbf{strawman} and \textbf{red herring} are both fallacies of relevance, but they involve slightly different shifts in the argument.  What is the main difference between these two types of fallacy? (\textbf{2 points})

\end{document}
