\documentclass[10pt]{article}

\usepackage[letterpaper]{geometry}
\usepackage{enumerate,verbatim,mdwlist}
\usepackage{fancyhdr}
\usepackage{linguex}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}

%Packages needed for trees
\usepackage{amsfonts,amsmath,amssymb}
\usepackage[varg]{txfonts}
\usepackage{qtree}
\usepackage{multicol}

%% Margin Setting
\geometry{hmargin={.5in,.5in},vmargin={1in,1in}}
\setlength{\parindent}{0.0in}
\setlength{\parskip}{3mm}
\setlength{\tabcolsep}{10pt}
\setlength{\arraycolsep}{10pt}

%% Header
\setlength{\headheight}{20.5pt}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[L]{\small Phil 101, f14}
\fancyhead[C]{\small Homework \# 5}
\fancyhead[R]{\small Due: Monday November 24, 2014 \\ Point total: 20 points}
\fancyfoot{}

%Compound events
\newcommand\negate[1]{\mathop{\mbox{$not$-$#1$}}}
\newcommand\disjoin[2]{\mathop{\mbox{$#1\; or\; #2$}}}
\newcommand\conjoin[2]{\mathop{\mbox{$#1\; and\; #2$}}}
\newcommand\given[2]{\mathop{\mbox{$#1\; given\; #2$}}}

\begin{document}

\textbf{Name:}\underline{  \textit{Answer Key}  }

\paragraph{Argument from analogy:} Consider the following argument from analogy.  For each of the facts that follow, first state what \textit{element} of the analogy the fact pertains to \textbf{(1 point each)}. The elements of every analogy are: \textit{primary analogue, secondary analogue, similarity, difference}, and the \textit{proposed property}. Based on the role the fact plays in the analogy, state whether the fact makes the analogy \textit{stronger}, \textit{weaker}, or whether it \textit{makes no difference} \textbf{(1 point each)}.

\begin{quote}
 Harold needs to have his rugs cleaned, and his friend Veronica reports that Ajax Carpet Service did an excellent job on her rugs. From this, Harold concludes that Ajax will do an equally good job on his rugs.
\end{quote}

\begin{enumerate} \itemsep7mm

\item Veronica hired Ajax several times, and Ajax always did an excellent job.

\textit{Each time rugs have gone to Ajax represents an additional primary analogue.  Since Ajax did an excellent job each of these times, this fact provides more evidence that Ajax consistently does good work.  Thus, the argument is strengthened.}

\item Veronica always had her rugs cleaned in mid-October, whereas Harold wants his done a week before Easter.

\textit{This fact represents a difference between the primary and secondary analogues.  Generally, extra differences weaken an analogy, but in this instance, the time of year doesn't seem all that relevant to whether the rugs will get clean.  So, this fact probably makes no difference to the analogy.}

\item Veronica's rugs had a few big stains on them before they were cleaned, and Harold's have only minor stains.

\textit{This fact also represents a difference between the primary and secondary analogues.  However, it is a difference that actually makes the analogy stronger. Since Harold's rugs are less dirty than Veronica's, it should be easier for Ajax to get them just as clean.}

\item Harold becomes less strict, and only demands that Ajax get his rugs approximately as clean as they got Veronica's.

\textit{This fact changes the proposed property that Harold is positing of his rugs.  Since he is being less specific about their cleanliness, it is easier for Ajax to meet the standard, and the analogy is made stronger.}

\item Harold reads 6 reviews of Ajax on Yelp, and all of them were very pleased with the service.

\textit{These reviews represent additional primary analogues.  Since each additional review was positive, this suggests that it is more likely that Harold's experience will also be positive.  So, the analogy is made stronger.}

\suspend{enumerate}

\paragraph{Theory of probability:} Consider the probability values: 

\[P(A) = 1/3 \hspace{1cm} P(B) = 3/5  \hspace{1cm} P(A\; given\; B) = 3/7 \hspace{1cm} P(A\; given\; not-B) = 0\]

\resume{enumerate} \itemsep7mm

 \item What does the fact that $P(A\; given\; B)$ is not zero mean about the relation between events $A$ and $B$? \textbf{(2 points)}
 
 \textit{This means that A is dependent on B.}
 
 \item Calculate $P(A\; and\; B)$ \textbf{(2 points)}:
 
 (To get this, I used the value calculated for $P(\given{B}{A})$ in \#9.)
 
 \[P(\conjoin{A}{B}) = P(A) \times P(\given{B}{A}) = \frac{1}{3} \times \frac{27}{35} = \frac{9}{35} \]
 
 \item Calculate $P(A\; or\; not-B)$ \textbf{(2 points)}:
 
 \[P(\disjoin{A}{\negate{B}}) = P(A) + P(\negate{B}) = P(A) + (1 - P(B)) = \frac{1}{3} + 1 - \frac{3}{5} = \frac{11}{15}\]
 
 \item Calculate $P(B\; given\; A)$ (hint: use Bayes' rule, \textbf{2 points}):
 
 \[P(\given{B}{A}) = \frac{P(B) \times P(\given{A}{B})}{P(A)} = \frac{\frac{3}{5} \times \frac{3}{7}}{\frac{1}{3}} = \frac{\frac{9}{35}}{\frac{1}{3}} = \frac{9}{35} \times \frac{3}{1} = \frac{27}{35} \]
 
 \item You devise an experiment to test hypothesis $H$. Prior to investigating more deeply, you believe $H$ to have a probability of $0.6$. The possible results of your experiment are evidence $E_1$, $E_2$, and $E_3$.  The probability that the result will be $E_1$ is $0.8$. And the likelihood of $E_1$ given $H$ is $0.75$.  You perform the experiment and the result is in fact $E_1$. Based on this information, use Bayes' rule to calculate the final probability of $H$ based on your experiment. \textbf{(2 points)}
 
 \[P(\given{H}{E_1}) = \frac{P(H)\times P(\given{E_1}{H})}{P(E_1)} = \frac{0.6 \times 0.75}{0.8} = 0.5625\]
\end{enumerate}
  
\paragraph{Bonus:}  Explain what goes wrong when a person fails to include the \textbf{base rate} in their calculation of the probability of a hypothesis. \textbf{(3 points)}

\textit{When we try to determine the overall probability of a hypothesis, we can't rely only on how much the evidence supports the the hypothesis.  We also have to factor in the initial plausibility of the hypothesis.  If the hypothesis is wildly implausible to start, then even if the evidence supports it pretty well, it may still be quite improbable overall. Ignoring the initial plausibility of the hypothesis is known as the base rate fallacy.}

\end{document}
