\documentclass[10pt]{article}

\usepackage[letterpaper]{geometry}
\usepackage{enumerate,verbatim,mdwlist}
\usepackage{fancyhdr}
\usepackage{linguex}
\usepackage[colorlinks=true,linkcolor=blue]{hyperref}

%Packages needed for trees
\usepackage{amsfonts,amsmath,amssymb}
\usepackage[varg]{txfonts}
\usepackage{qtree}
\usepackage{multicol}

%% Margin Setting
\geometry{hmargin={.5in,.5in},vmargin={1in,1in}}
\setlength{\parindent}{0.0in}
\setlength{\parskip}{3mm}
\setlength{\tabcolsep}{10pt}
\setlength{\arraycolsep}{10pt}

%% Header
\setlength{\headheight}{20.5pt}
\pagestyle{fancy}
\fancyhead{}
\fancyhead[L]{\small Phil 101, f14}
\fancyhead[C]{\small Homework \# 6}
\fancyhead[R]{\small Due: Wednesday December 10, 2014 \\ Point total: 20 points}
\fancyfoot{}

\begin{document}

\textbf{Name:}\underline{\hspace{2in}}

\paragraph{Construct an argument:} The four claims below are sometimes labeled \textit{pseudoscientific}, which is meant to disparage them as unsupported by acceptable scientific techniques.  But for each of the claims, there are many people who argue adamently in its favor. Your task is to select one of the claims below and argue either in favor of it or against it.

\paragraph{The claims you can choose from:}
\begin{itemize}
 \item A person's astrological sign can be used to predict their fortune.
 \item Intelligent extra terrestrials have visited the Earth.
 \item Homeopathic medicine can be helpful in treating certain diseases.
 \item The world was created by God around 6,000 years ago.
\end{itemize}

\paragraph{Some guidelines:} This assignment is meant as a fun exercise for you to apply some of what we've learned in this class to real life debates.  In arguing for your position, please follow these guidelines:
\begin{itemize}
 \item Your argument should be \textbf{at least 2 paragraphs long}, but \textbf{no longer than 1 full page}.
 \item Please type your argument to make it easier to read for grading purposes.
 \item Clearly state which claim you are addressing, and what stance (positive or negative) you are taking on it.
 \item Explain the main premises in your argument as well as the reasoning you use to support your position.
 \item Try to involve some of the ideas we have explored in this class in explaining your argument. (Ch. 14 of the book may be particularly helpful.)
 \item You may want to do a little outside research to familiarize yourself with the claim you choose before jumping right into your argument.  Wikipedia is an acceptable source for this project.
 \item You will be graded on the \textit{clarity} of your argument, your \textit{ability to address the claim} under consideration, and how well you \textit{apply ideas from this class}.
 \item You will \textbf{not} be graded on \textit{grammar}  or \textit{writing style} as long as your argument is readable.
 \item I do not expect you to produce an entirely novel way of thinking about the claim you choose. I want to see your own thinking on the subject, but you don't need to come up with something no one has ever thought of before.
\end{itemize}


\paragraph{Bonus:}  Consider one response that an opponent might make to your argument, and defend your argument against that response. \textbf{(5 points)}

\end{document}
