\documentclass[letterpaper,12pt]{article}

\usepackage[colorlinks=true,linkcolor=blue]{hyperref}
\usepackage{amssymb}

%Margin settings
\usepackage{geometry}
\geometry{hmargin={1in,1in},vmargin={1in,1in}}

%Header settings
\usepackage{fancyhdr}
\setlength{\headheight}{15pt}
\pagestyle{fancy}
\lhead{Phil 101, f14}
\chead{Recognizing fallacies}
\rhead{September 24, 2014}

%\pagenumbering{}

%Paragraph settings
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex plus .5ex minus .2ex}

\begin{document}

\paragraph{Directions:} For each of the passages below, determine what fallacy, if any, the argument commits.  All of the fallacies are either fallacies of \textbf{relevance} or fallacies of \textbf{weak induction}.  Some of the passages may not commit a fallacy at all.

\begin{enumerate}
  \setlength\itemsep{20pt}
  
  \item ``Of course Craig says that ethanol is a good fuel additive.  He's a country farm boy who grew up with corn growing in his front yard.  I don't think we need to bother with his energy policy.''
  
  \item ``Each of the last three weekends, Martin has gone to sleep with his shoes on.  He's also woken up each of the next mornings feeling nauseous with a pounding head ache.  So, if you don't want to get sick, always take your shoes off before going to bed.''
  
  \item ``Walt claims that immigrants should have a path to citizenship.  But jobs are really important.  Without them, we can't grow the economy, and crime rates will grow.  Clearly, Walt needs to rethink his position.''
  
  \item ``Zo\"{e} got really sick after dinner last night.  We had all the same stuff, except Zo\"{e} tried mussels for the first time.  I guess she should stay away from seafood in the future.''
  
  \item ``Every time I go to Old Queens Tavern, I get accosted by a drunk jerk in a Scarlet Knights shirt.  Rutgers sports fans are the worst!''
  
  \item ``Am I happy with how my project turned out? Well, first I had to fix everyone's breakfast before working from 9 until 6.  Then I was all over town with errands. Finally, I made dinner and got everyone prepared for tomorrow. But, yeah, I got it done.''
  
\end{enumerate}


\end{document}
